### 2022-11-13 NSF2022/Description/backgroundv3.tex
yny 1113a

When fluids are involved,
tracking self-organization
becomes difficult due to the deformability of
suspended media.  A dominant approach in this area,
the phase field method
introduce a scalar-valued phase
field function for the volume fraction of
amphiphilic lipids or polymers in solvent, for example.
Sometimes called diffuse interface
methods, one devises energy functionals involving the gradient and
well-potentials of the phase field, and possibly long range interaction
kernels, 
\cite{Promislow2022UndulatedBI,C9SM01983A,doi:10.1063/5.0009734,
LiAn-Chang16,Choksi2003OnTD}.
Because the fluid interfaces are given by the
level-set of a smooth function,
they can have almost arbitrary topology
e.g., free ends, junctions
\cite{Promislow2017ExistenceBA,Promislow2022UndulatedBI},
the evolution equations can be studied
by well-established tools variational and analytical means,
\cite{Gavish2011CurvatureDF,Dai2019WeakSF,Dai2015CompetitiveGE,
Dai2022GeometricEO,Dai2020MinimizersFT,Dai2013GeometricEO},
and the governing equations are amenable to numerical approximation
\cite{Christlieb2020BenchmarkCO,Christlieb2019CompetitionAC}.
One of the downsides, though, 
is that granular details like molecular orientations and
shape changes induced by proteins that may be crucial to the physics, are
clumsy to incorporate in the model. Moreover, numerical simulations of
phase field models typically discretize the entire fluid domain under
consideration, and apply cutoff boundary conditions when periodic
domains are not under consideration.

@@

A striking aspect of our hydrodynamic
simulation is its fidelity to physics.
The model contains a few
parameters; granule radius $c$,
screening length $\epsilon$,
surface tension $\gamma$.
When set to realistic values,
\cite{Fu2018_SIAM, ErLjCl89, Lin2005, Parsegian, Israelachvili80, GarciaSaez, KUZMIN2005, Petelska2012,Jackson2016},
we recover many of the well-known
elastic moduli for single-component lipid bilayers
\cite{Nagle17, Nagle17-2, LeVeWa14,NAGLE2000159}.

@@

In the absence of external flow, the
system~\eqref{eq:RBT}--\eqref{eqn:stressbalance} has the energy law
\begin{align}
\label{eq:energy_law}
  \frac{d}{dt}E
  = - \int_{\Omega(t)}\tfrac{1}{2}\mu|\nabla \uu + \nabla
  \uu^T|^2 \,d\xx,\quad
    E = \gamma \int_{\Omega(t)}
  \frac{\epsilon}{2} |\nabla \phi|^2 + \frac{1}{\epsilon} W(\phi) \,d\xx.
\end{align}
Here $E$ is the energy stored in the phase transitions
and the energy law states that
the granules move to decrease energy,
which is lost due to viscous dissipation.
The derivation of \eqref{eq:energy_law} is interesting
because $E$ depends both on $\phi$
and $\Omega(t)$. It uses the Rayleigh Transport Theorem to write the time
derivative of $E$ as an energy density flux through the boundary
involving the phase stress $T_i$, and a bulk term involving the Euler-Lagrange
derivative of $E$~\cite{Fu2018_SIAM}. The Euler-Lagrange derivative vanishes due
to~\eqref{eqn:phase} and the boundary flux transforms into viscous
dissipation due to~\eqref{eqn:hydro_stress} and the Stokes equations.

\begin{wrapfigure}[16]{l}{4in} 
\includegraphics[width=4in]{figures/Background/coarsening.pdf}
\vspace{-20pt}
\caption{\label{fig:coarsening}
Modeling procedure.
We initilize a suspension of granules.
Under the moving domain problem
\eqref{eq:RBT}-\eqref{eqn:hydro_stress}
with amphiphilic boundary condition, the granules
self-organize into vesicle bilayers.
We then introduce external hydrodynamic effects to study elastic properties.
}
\vspace{5pt}
\end{wrapfigure}
In the far field, $\lim_{\xx \to \infty} \phi(\xx) = \phi_0$ and for the
background flow, we require that $\lim_{\xx \to \infty} \uu(\xx) -
\uu_{\infty}(\xx) = \mathbf{0}$, where
the background flow takes the form
%In the case of shear background flow,
%for example,
%\begin{align}
%\label{eqn:shear_BG_flow}
%\uu_{\infty}(\xx) = \dot{\gamma} \xx \cdot \mathbf{e}_y \mathbf{e}_x,
%\end{align}
%where $\dot \gamma$ is the shear rate and $\mathbf{e}_x$, $\mathbf{e}_y$
%are horizontal and vertical basis vectors, respectively.
%Other
%background flows include
shear, extensional, parabolic Poiseuille, and
Taylor-Green flow, for example.
For numerical stability and physical realism, we
include pairwise repulsive forces and torques
in~\eqref{eqn:stressbalance} to prevent the collision between
neighboring granules which introduces a repulsive potential in the energy
law~\eqref{eq:energy_law}. Background flow adds elastic energy to the
granule configurations.
%and is quantified by simulation. We refer to the
%PIs work~\cite{FuQuRyYo22, fu-ryh-qua-you2022, Fu2018_SIAM} for details.

@@

\subsection{Outline of the proposal}
The aim of this proposal is to advance the simulation
and theory for soft-matter systems
and foster training in mathematical sciences.
Section \ref{sec:BroaderImpacts} summarizes the broader
impacts and intellectual merit of our research activity.
Section \ref{sec:proposed-work} contains our
three specific aims.  They are
mathematical analysis of the system
(\S \ref{sec:specificaim1}: PI Ryham),
numerical implementation
(\S \ref{sec:specificaim2}: PI Quaife),
and  kinetic theory
(\S \ref{sec:specificaim3}: PI Young).

>>>>

When fluids are involved,
tracking self-organization
becomes difficult due to the deformability of
suspended media in the presence of non-local hydrodynamic interactions.  
A dominant approach in this area,
the phase field method
introduce a scalar-valued phase
field function for the volume fraction of
amphiphilic lipids or polymers in solvent, for example.
Sometimes called diffuse interface
methods, one devises energy functionals involving the gradient and
well-potentials of the phase field, and possibly long range interaction
kernels, 
\cite{Promislow2022UndulatedBI,C9SM01983A,doi:10.1063/5.0009734,
LiAn-Chang16,Choksi2003OnTD}.
Because the fluid interfaces are given by the
level-set of a smooth function,
they can have almost arbitrary topology
e.g., free ends, junctions
\cite{Promislow2017ExistenceBA,Promislow2022UndulatedBI},
the evolution equations can be studied
by well-established tools variational and analytical means,
\cite{Gavish2011CurvatureDF,Dai2019WeakSF,Dai2015CompetitiveGE,
Dai2022GeometricEO,Dai2020MinimizersFT,Dai2013GeometricEO},
and the governing equations are amenable to numerical approximation
\cite{Christlieb2020BenchmarkCO,Christlieb2019CompetitionAC}.
One of the downsides, though, 
is that granular details like molecular orientations and
shape changes induced by proteins that may be crucial to the physics, are
clumsy to incorporate in the model. Moreover, numerical simulations of
phase field models typically discretize the entire fluid domain under
consideration, and apply cutoff boundary conditions when periodic
domains are not under consideration.

@@

A striking aspect of our hydrodynamic
simulation is its fidelity to physics.
The model contains a few
parameters; granule radius $c$,
screening length $\epsilon$,
surface tension $\gamma$.
When set to realistic values in the literature
\cite{Fu2018_SIAM, ErLjCl89, Lin2005, Parsegian, Israelachvili80, GarciaSaez, KUZMIN2005, Petelska2012,Jackson2016},
we recover many of the well-known
elastic moduli for single-component lipid bilayers
\cite{Nagle17, Nagle17-2, LeVeWa14,NAGLE2000159}.

@@

In the absence of external flow, the
system~\eqref{eq:RBT}--\eqref{eqn:stressbalance} has the energy law
\begin{align}
\label{eq:energy_law}
  \frac{d}{dt}E
  = - \int_{\Omega(t)}\tfrac{1}{2}\mu|\nabla \uu + \nabla
  \uu^T|^2 \,d\xx,\quad
    E = \gamma \int_{\Omega(t)}
  \frac{\epsilon}{2} |\nabla \phi|^2 + \frac{1}{\epsilon} W(\phi) \,d\xx.
\end{align}
Here $E$ is the energy stored in the phase transitions
and the energy law states that
the granules move to decrease energy,
which is lost due to viscous dissipation.
The derivation of \eqref{eq:energy_law} is interesting
because $E$ depends both on $\phi$
and $\Omega(t)$. It uses the Rayleigh Transport Theorem to write the time
derivative of $E$ as an energy density flux through the boundary
involving the phase stress $T_i$, and a bulk term involving the Euler-Lagrange
derivative of $E$~\cite{Fu2018_SIAM}. The Euler-Lagrange derivative vanishes due
to~\eqref{eqn:phase} and the boundary flux transforms into viscous
dissipation due to~\eqref{eqn:hydro_stress} and the Stokes equations. Figure~\ref{fig:coarsening} summarizes the general evolution of $E$ from the self-assembly process to
the collective dynamics of the assembly under external forcing.

\begin{wrapfigure}[17]{l}{4in} 
\includegraphics[width=4in]{figures/Background/coarsening.pdf}
\vspace{-20pt}
\caption{\label{fig:coarsening}
Energy $E$ in equation~(\ref{eq:energy_law}) of a suspension of amphiphilic granules versus time $t$.
Under the moving domain problem
\eqref{eq:RBT}-\eqref{eqn:hydro_stress}
with amphiphilic boundary condition, the granules
self-organize into vesicle bilayers.
External hydrodynamic effects are introduced to study elastic properties the the granules assembly \cite{Fu2018_SIAM,FuQuRyYo22}.
}
\vspace{5pt}
\end{wrapfigure}
In the far field, $\lim_{\xx \to \infty} \phi(\xx) = \phi_0$ and for the
background flow, we require that $\lim_{\xx \to \infty} \uu(\xx) -
\uu_{\infty}(\xx) = \mathbf{0}$, where
the background flow takes the form
%In the case of shear background flow,
%for example,
%\begin{align}
%\label{eqn:shear_BG_flow}
%\uu_{\infty}(\xx) = \dot{\gamma} \xx \cdot \mathbf{e}_y \mathbf{e}_x,
%\end{align}
%where $\dot \gamma$ is the shear rate and $\mathbf{e}_x$, $\mathbf{e}_y$
%are horizontal and vertical basis vectors, respectively.
%Other
%background flows include
shear, extensional, parabolic Poiseuille, and
Taylor-Green flow, for example.
For numerical stability and physical realism, we
include pairwise repulsive forces and torques
in~\eqref{eqn:stressbalance} to prevent the collision between
neighboring granules which introduces a repulsive potential in the energy
law~\eqref{eq:energy_law}. Background flow adds elastic energy to the
granule configurations.
%and is quantified by simulation. We refer to the
%PIs work~\cite{FuQuRyYo22, fu-ryh-qua-you2022, Fu2018_SIAM} for details.

@@

\subsection{Outline of the proposal}
The aim of this proposal is to advance the simulation
and theory for soft-matter systems
and foster training in mathematical sciences.
Section \ref{sec:BroaderImpacts} summarizes the broader
impacts and section \ref{sec:IntellectualMerit} summarizes the intellectual merit of our research activity.
Section \ref{sec:proposed-work} contains our
three specific aims.  They are
mathematical analysis of the system
(\S \ref{sec:specificaim1}: PI Ryham),
numerical implementation
(\S \ref{sec:specificaim2}: PI Quaife),
and  kinetic theory
(\S \ref{sec:specificaim3}: PI Young).

### 2022-11-13 NSF2022/Description/broaderimpacts.tex
yny 1113a

\section{Intellectual Merit}
The purpose of this research is to reach interesting physical phenomena with less computational cost than molecular dynamics, and account for more general features that continuum theory misses. The main ingredient is defining a nonlocal interaction through the solution of an elliptic boundary value problem that has the phenomenological characteristics of long-range hydrophobic attraction. This minimal model, while intuitive, is quite a general description of amphiphiles in solvent and gives rise to rich phenomena from Janus particle aggregates to correctly predicting elastic properties of bilayer. The technical research tasks include quantifying collective properties of amphiphilic ensembles, improving on mathematical models, efficient, high-order numerical algorithms for large-scale simulations, and incorporating external fields through electric charge.

>>>>

\section{Intellectual Merit}
\label{sec:IntellectualMerit}
The purpose of this research is to reach interesting physical phenomena with less computational cost than molecular dynamics, and account for more general features that continuum theory misses. The main ingredient is defining a nonlocal interaction through the solution of an elliptic boundary value problem that has the phenomenological characteristics of long-range hydrophobic attraction. This minimal model, while intuitive, is quite a general description of amphiphiles in solvent and gives rise to rich phenomena from Janus particle aggregates to correctly predicting elastic properties of bilayer. The technical research tasks include quantifying collective properties of amphiphilic ensembles, improving on mathematical models, efficient, high-order numerical algorithms for large-scale simulations, and incorporating external fields through electric charge.
